plyranges: a grammar of genomic data transformation

### 2018-05-22 paper.Rmd
add more detailed author affiliations

---
title: "plyranges: a grammar for transforming genomics data" 
author:
  - name: Stuart Lee
    email: stuart.lee1@monash.edu
    affiliation: Monash University
  - name: Michael Lawrence
    email: michael.lawrence@gene.com
    affiliation: Genentech
  - name: Di Cook
    email: di.cook@monash.edu
    affiliation: Monash University
abstract: "The Bioconductor project has created many useful data abstractions for analysing high-throughput genomics experiments. However, there is a cognitive load placed on a user in learning a data abstraction and understanding its appropriate use. Throughout a standard workflow, a user must navigate and know many of these abstractions to perform an genomic analysis task, when a single data abstraction, a GRanges object will suffice. The GRanges class naturally represent genomic intervals and their associated measurements. By recognising that the GRanges class follows 'tidy' data principles we have created a grammar of genomic data transformation. The grammar defines verbs for performing actions on and between genomic interval data. It provides a principled way of performing common genomic data analysis tasks through a coherent interface to existing Bioconductor infrastructure, resulting in human readable analysis workflows. We have implemented this grammar as a Bioconductor/R package called plyranges."
keywords: "Bioconductor, Grammar, Genomes, Data Analysis"
bibliography: references.bib
output: 
  pdf_document:
    template: svm-latex-ms.tex
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
    citation_package: biblatex
fontfamily: mathpazo
---

```{r setup, include = FALSE}
library(tibble)
library(knitr)
opts_chunk$set(message = FALSE, 
               warning = FALSE, 
               comment = "#>", 
               fig.width = 5, 
               fig.height = 3, 
               fig.align = "center")
# for peak figures
library(ggbio)

>>>>

---
title: "plyranges: a grammar of genomic data transformation" 
author:
  - name: Stuart Lee
    email: stuart.lee1@monash.edu
    affiliation: |
     | Department of Econometrics and Business Statistics, Monash University
     |                        Molecular Medicine Division, Walter and Eliza Hall Institute
  - name: Di Cook
    email: di.cook@monash.edu
    affiliation: Department of Econometrics and Business Statistics, Monash University
  - name: Michael Lawrence
    email: michael.lawrence@gene.com
    affiliation: Bioinformatics and Computational Biology, Genentech, Inc.
abstract: "The Bioconductor project has created many useful data abstractions for analysing high-throughput genomics experiments. However, there is a cognitive load placed on a user in learning a data abstraction and understanding its appropriate use. Throughout a standard workflow, a user must navigate and know many of these abstractions to perform an genomic analysis task, when a single data abstraction, a GRanges object will suffice. The GRanges class naturally represent genomic intervals and their associated measurements. By recognising that the GRanges class follows 'tidy' data principles we have created a grammar of genomic data transformation. The grammar defines verbs for performing actions on and between genomic interval data. It provides a principled way of performing common genomic data analysis tasks through a coherent interface to existing Bioconductor infrastructure, resulting in human readable analysis workflows. We have implemented this grammar as a Bioconductor/R package called \\texttt{plyranges}."
keywords: "Bioconductor, Grammar, Genomes, Data Analysis"
bibliography: references.bib
output: 
  pdf_document:
    template: svm-latex-ms.tex
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
    citation_package: biblatex
fontfamily: mathpazo
---
```{r setup, include = FALSE}
library(tibble)
library(knitr)
opts_chunk$set(message = FALSE, 
               warning = FALSE, 
               comment = "#>", 
               fig.width = 5, 
               fig.height = 3, 
               fig.align = "center")
# for peak figures
library(ggbio)

### 2018-05-22 paper.tex
add more detailed author affiliations

\title{plyranges: a grammar for transforming genomics data  }

\author{\Large Stuart Lee\vspace{0.05in} \newline\normalsize\emph{Monash University}   \and \Large Michael Lawrence\vspace{0.05in} \newline\normalsize\emph{Genentech}   \and \Large Di Cook\vspace{0.05in} \newline\normalsize\emph{Monash University}  }

@@

\@ifpackageloaded{color}{
    \PassOptionsToPackage{usenames,dvipsnames}{color}
}{%
    \usepackage[usenames,dvipsnames]{color}
}
\makeatother
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={Stuart Lee (Monash University) and Michael Lawrence (Genentech) and Di Cook (Monash University)},
             pdfkeywords = {Bioconductor, Grammar, Genomes, Data Analysis},  
            pdftitle={plyranges: a grammar for transforming genomics data},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

@@

{
   \vskip 13.5pt\relax \normalsize\fontsize{11}{12} 
\textbf{\authorfont Stuart Lee} \hskip 15pt \emph{\small Monash University}   \par \textbf{\authorfont Michael Lawrence} \hskip 15pt \emph{\small Genentech}   \par \textbf{\authorfont Di Cook} \hskip 15pt \emph{\small Monash University}

@@

\noindent The Bioconductor project has created many useful data abstractions for
analysing high-throughput genomics experiments. However, there is a
cognitive load placed on a user in learning a data abstraction and
understanding its appropriate use. Throughout a standard workflow, a
user must navigate and know many of these abstractions to perform an
genomic analysis task, when a single data abstraction, a GRanges object
will suffice. The GRanges class naturally represent genomic intervals
and their associated measurements. By recognising that the GRanges class
follows `tidy' data principles we have created a grammar of genomic data
transformation. The grammar defines verbs for performing actions on and
between genomic interval data. It provides a principled way of
performing common genomic data analysis tasks through a coherent
interface to existing Bioconductor infrastructure, resulting in human
readable analysis workflows. We have implemented this grammar as a
Bioconductor/R package called plyranges.

@@

\noindent  \section{Introduction}\label{introduction}

@@

\section{Genomic relational algebra}\label{genomic-relational-algebra}

\subsection{Data model}\label{data-model}

@@

\subsection{Algebraic operations}\label{algebraic-operations}

@@

\subsection{Design principles}\label{design-principles}

@@

\subsubsection{Cognitive consistency and
fluency}\label{cognitive-consistency-and-fluency}

@@

\subsubsection{Cohesion}\label{cohesion}

@@

\subsubsection{Expressiveness}\label{expressiveness}

@@

\section{Results}\label{results}

@@

\subsection{Peak Finding}\label{peak-finding}

@@

\subsection{Computing Windowed
Statistics}\label{computing-windowed-statistics}

@@

\subsection{Quality Control Metrics}\label{quality-control-metrics}

@@

\section{Discussion and conclusion}\label{discussion-and-conclusion}

>>>>

\title{plyranges: a grammar of genomic data transformation  }

\author{\Large Stuart Lee\vspace{0.05in} \newline\normalsize\emph{Department of Econometrics and Business Statistics, Monash University\\
\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}Molecular
Medicine Division, Walter and Eliza Hall Institute}   \and \Large Di Cook\vspace{0.05in} \newline\normalsize\emph{Department of Econometrics and Business Statistics, Monash University}   \and \Large Michael Lawrence\vspace{0.05in} \newline\normalsize\emph{Bioinformatics and Computational Biology, Genentech, Inc.}  }

@@

\@ifpackageloaded{color}{
    \PassOptionsToPackage{usenames,dvipsnames}{color}
}{%
    \usepackage[usenames,dvipsnames]{color}
}
\makeatother
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={Stuart Lee (Department of Econometrics and Business Statistics, Monash University\\
\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}Molecular
Medicine Division, Walter and Eliza Hall Institute) and Di Cook (Department of Econometrics and Business Statistics, Monash University) and Michael Lawrence (Bioinformatics and Computational Biology, Genentech, Inc.)},
             pdfkeywords = {Bioconductor, Grammar, Genomes, Data Analysis},  
            pdftitle={plyranges: a grammar of genomic data transformation},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

@@

{
   \vskip 13.5pt\relax \normalsize\fontsize{11}{12} 
\textbf{\authorfont Stuart Lee} \hskip 15pt \emph{\small Department of Econometrics and Business Statistics, Monash University\\
\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}Molecular
Medicine Division, Walter and Eliza Hall Institute}   \par \textbf{\authorfont Di Cook} \hskip 15pt \emph{\small Department of Econometrics and Business Statistics, Monash University}   \par \textbf{\authorfont Michael Lawrence} \hskip 15pt \emph{\small Bioinformatics and Computational Biology, Genentech, Inc.}

@@

\noindent The Bioconductor project has created many useful data abstractions for
analysing high-throughput genomics experiments. However, there is a
cognitive load placed on a user in learning a data abstraction and
understanding its appropriate use. Throughout a standard workflow, a
user must navigate and know many of these abstractions to perform an
genomic analysis task, when a single data abstraction, a GRanges object
will suffice. The GRanges class naturally represent genomic intervals
and their associated measurements. By recognising that the GRanges class
follows `tidy' data principles we have created a grammar of genomic data
transformation. The grammar defines verbs for performing actions on and
between genomic interval data. It provides a principled way of
performing common genomic data analysis tasks through a coherent
interface to existing Bioconductor infrastructure, resulting in human
readable analysis workflows. We have implemented this grammar as a
Bioconductor/R package called \texttt{plyranges}.

@@

\noindent  \hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

@@

\hypertarget{genomic-relational-algebra}{%
\section{Genomic relational algebra}\label{genomic-relational-algebra}}

\hypertarget{data-model}{%
\subsection{Data model}\label{data-model}}

@@

\hypertarget{algebraic-operations}{%
\subsection{Algebraic operations}\label{algebraic-operations}}

@@

\hypertarget{design-principles}{%
\subsection{Design principles}\label{design-principles}}

@@

\hypertarget{cognitive-consistency-and-fluency}{%
\subsubsection{Cognitive consistency and
fluency}\label{cognitive-consistency-and-fluency}}

@@

\hypertarget{cohesion}{%
\subsubsection{Cohesion}\label{cohesion}}

@@

\hypertarget{expressiveness}{%
\subsubsection{Expressiveness}\label{expressiveness}}

@@

\hypertarget{results}{%
\section{Results}\label{results}}

@@

\hypertarget{peak-finding}{%
\subsection{Peak Finding}\label{peak-finding}}

@@

\hypertarget{computing-windowed-statistics}{%
\subsection{Computing Windowed
Statistics}\label{computing-windowed-statistics}}

@@

\hypertarget{quality-control-metrics}{%
\subsection{Quality Control Metrics}\label{quality-control-metrics}}

@@

\hypertarget{discussion-and-conclusion}{%
\section{Discussion and conclusion}\label{discussion-and-conclusion}}
